---
title: "Hypertension_Simulation"
author: "Hemanshu Das"
date: "5/21/2019"
output: html_document
---

The Rmarkdown is an implementation of the simulation model to emulate a cohort of 50 yr old, Male, 27.7BMI Smokers with 150mmHg SBP profile through state-level transitions with constant death rate for a period of 10 years. </br>

The intervention considered is simplistic wherein everyone in the treatment cohort has access to hypertension treatment. 

```{r, echo=FALSE}
rm(list = ls())               #removing any variables in R's memory
options(max.print=10000)
library(ggplot2)
start_time <- Sys.time()
```
```{r, echo=TRUE}
#Model Inputs
n.i <- 1000                            #Number of individuals
ini_age <- 50                           #Start Age of Cohort
end_age <- 60                           #Simulation End Age
cl <- 1                                 #Number of Cycle in a year
v.n <- c("W","MI","ST","postMI","postST","cD","D")        #Model States: W-Well, CVD-CVD Event, MI - Myocardial Infraction Transitory State, ST- Transitory Stroke State, cD - CVD Death, D - Death due to other causes
n.s <- length(v.n)                      #Number of model states
v.M_1 <- rep("W",n.i)                   #Everyone is Well at the start
v.Str <- c("No Treatment","Treatment")  #Storing the Strategy Indicator
n.t <- (end_age - ini_age)*12                #Simulation cycles
d.c   <- d.e <- 0.03/12                    #Equal discounting of costs and QALYs by 3% p.a

#Initialization of Transition Probabilities (Monthly transition)
p.W_CVD <- 0.00206                  #Probability of CVD transitory event when well; 50 yr, M, Smoker, SBP 150mmHg, BMI 27.7
p.CVD_MI <- 0.188                  #Probability of MI when CVD event occurs
p.CVD_ST <- 1 - p.CVD_MI           #Probability of Stroke when CVD event occurs
p.MI_cD <- 0.335                   #Probability of fatal outcome when MI occurs
p.ST_cD <- 0.181                   #Probability of fatal outcome when Stroke occurs
p.MI_postMI <- 1 - p.MI_cD         #Probability of non-fatal outcome when MI occurs
p.ST_postST <- 1 - p.ST_cD         #Probability of non-fatal outcome when Stroke occurs
p.postMI_cD <- 0.0022*0.76              #Probability of CVD death post non-fatal MI
p.postMI_D <- 0.0022*0.24              #Probability of death post non-fatal MI
p.postST_cD <- 0.0028 *0.336             #Probability of CVD death post non-fatal ST
p.postST_D <- 0.0028 *0.664             #Probability of death post non-fatal ST
p.W_D <- 0.00054                   #Probability of normal death from any state

#Intiatlization of Cost for each state
c.W <- 0                      #cost of being well
c.MI <- 1428.57                  #cost of acute MI treatment (Assumed to be INR 1,00,000)
c.ST <- 2142.85                  #cost of acute Stroke treatment (Assumed to be INR 1,50,000)
c.postMI <- c.postST <- 28.57 #cost of follow up care (Assumed to be INR 2000 p.m)
c.D <- c.cD <- 0              #cost of Death

#Intialization of Utility for each state
u.W <- 1                      #utility of being well
u.MI <- u.postMI <- 0.76                  #utility of MI state
u.ST <- u.postST <- 0.629                 #utility of Stroke state
u.D <- u.cD <- 0              #utility of death

# If not treated
p.W_CVD_NT <- 0.00206          #15% reduction in CVD risk
c.W_NT <- 0                   #Cost of medication and consultation in a month

# If treated
p.W_CVD_T <- 0.00172          #15% reduction in CVD risk
c.W_T <- 4.57                 #Cost of medication and consultation in a month

```

```{r, echo=FALSE}
####Function to simulate the markov cycle tree
MicroSim <- function(p.W_CVD,c.W,TS.out = TRUE,TR.out=TRUE,seed=1) { 
#Arguments:
          #TS.out:  Flag for matrix of transitions between states
          #TR.out:  Flag for microsimulation trace
          #p.UD:    Probability of death when unwell attributed to event
          #c.U:     Cost of staying unwell due to cost of drugs

          v.dwc <- 1 / (1 + d.c) ^ (0:n.t)            #Vector Multiplication for calculating costs
          v.dwu <- 1 / (1 + d.e) ^ (0:n.t)           #Vector Multiplication for calculating QALY

#Matrixes to capture state, cost and health outcomes for all individuals at any time point
m.M <- m.C <- m.U <- matrix(nrow = n.i, ncol=n.t+1,
                            dimnames = list(paste("ind",1:n.i,sep = " "),
                                            paste("cycle",0:n.t, sep = " ")))
m.M[,1] <- v.M_1              #Initial health state

for (i in 1:n.i) {
          seed=sample.int(100,1)
          set.seed(seed+i)
          m.C[i,1] <- Cost(m.M[i,1],c.W)              #Costs per individual for initial health state
          m.U[i,1] <- Util(m.M[i,1])              #QALY per individual for initial health state
          
          for (t in 1:n.t) {
                    v.p <- Probs(m.M[i,t],p.W_CVD)        #Transition probabilities at cycle t
                    
                    m.M[i, t+1] <- sample(v.n, prob = v.p, size =1)   #Sample the next health state
                    m.C[i, t+1] <- Cost(m.M[i,t+1],c.W)               #Costs per individual at cycle t+1
                    m.U[i, t+1] <- Util(m.M[i,t+1])                   #QALY per individual at cycle t+1
          }                                       #Closing loop for cycles
                                         #Closing loop for progress display
}

#Calculate number of CVD Deaths
count_cD <- 0
for (i in 1:n.i) {
          if ((m.M[i,n.t+1]) == "cD") {
                    count_cD = count_cD + 1
          }
}

#Calculate number of Non-CVD Deaths
count_D <- 0
for (i in 1:n.i) {
          if ((m.M[i,n.t+1]) == "D") {
                    count_D = count_D + 1
          }
}

tc <- m.C %*% v.dwc           #Calculating the Costs
tu <- m.U %*% v.dwu           #Calculating the QALYs

tc_hat <- mean(tc)            #Average Cost
tu_hat <- mean(tu)            #Average QALY

#Optional Matrix of Transitions between states
if (TS.out == TRUE) {
          TS <- paste(m.M, cbind(m.M[,-1],NA), sep = "->")            #Transitions from one state to another
          TS <- matrix(TS,nrow=n.i)
          rownames(TS) <- paste("Ind", 1:n.i, sep = " ")              #Naming the rows
          colnames(TS) <- paste("Cycle", 0:n.t, sep = " ")            #Naming the columns
} else {
          TS <- NULL
}

#Optional Output Trace
if (TR.out == TRUE) {
          TR <- t(apply(m.M,2,function(x) table(factor(x,levels=v.n, ordered = TRUE))))
          TR <- TR/n.i                                                #Creating a Distribution Trace
          rownames(TR) <- paste("Cycle", 0:n.t, sep = " ")            #Naming the rows
          colnames(TR) <- v.n                                         #Naming the columns
} else {
          TR <- NULL
}

#Calculating the number of CVD events
count_CVD <- 0
 
for (i in 1:n.i) {
          for (t in 1:n.t) {
                              if (TS[i,t] == "W->MI" || TS[i,t] == "W->ST") {
                              count_CVD = count_CVD + 1 
                              }
          }
}

#Finding the distribution of states across states
m.CS <- matrix(nrow = 3, ncol=121,
               dimnames = list(c("Well","CVD Death","Death"),
                               paste("Cycle",0:120,sep = " ")))
m.CS[,] <- 0

for (i in 1:n.i) {
          for (t in 0:n.t+1) {
                    if (m.M[i,t] == "W") {
                              m.CS[1,t] = m.CS[1,t] + 1
                    }
                    else if (m.M[i,t] == "cD") {
                              m.CS[2,t] = m.CS[2,t] + 1
                    }
                    else if (m.M[i,t] == "D") {
                              m.CS[3,t] = m.CS[3,t] + 1
                    }
          }
}


results <- list(m.M = m.M, m.C = m.C, m.U = m.U, count_cD = count_cD, count_D = count_D, tc_hat = tc_hat, tu_hat = tu_hat, tu=tu, TS = TS, TR = TR, count_CVD = count_CVD, m.CS = m.CS)
return(results)
}                             #End of the MicroSim Function

####Probability Function to update the transition probabilities of every cycle
Probs <- function(M.it,p.W_CVD) {
          #M.it:    Health state occupied by individual i at cycle t
          #p.UD:    Probability of death when unwell due to event
          
          v.p.it <- rep(NA,n.s)                                       #Vector of transition probabilities
          names(v.p.it) <- v.n                                        #Naming the vector
          
          v.p.it[M.it == "W"] <- c(1-(p.W_CVD+p.W_D),p.W_CVD*p.CVD_MI,p.W_CVD*p.CVD_ST,0,0,0,p.W_D) #tranistion probability when well
          v.p.it[M.it == "MI"] <- c(0,0,0,p.MI_postMI,0,p.MI_cD,0)                          #transition probability when in MI state
          v.p.it[M.it == "ST"] <- c(0,0,0,0,p.ST_postST,p.ST_cD,0)                          #transition probability when in Stroke state
          v.p.it[M.it == "postMI"] <- c(0,0,0,(1-(p.postMI_cD+p.postMI_D)),0,p.postMI_cD,p.postMI_D)          #transition probability when in MI state
          v.p.it[M.it == "postST"] <- c(0,0,0,0,(1-(p.postST_cD+p.postST_D)),p.postST_cD,p.postST_D)                          #transition probability when in Stroke state
          v.p.it[M.it == "cD"] <- c(0,0,0,0,0,1,0)                     #transition probability when in CVD Death state
          v.p.it[M.it == "D"] <- c(0,0,0,0,0,0,1)                       #transition probability when in Death state
          return(v.p.it)                                              #returning probabilities
}

####Cost function estimates the cost at every cycle
Cost <- function (M.it, c.W) {
          #M.it:    Heath state occupied by individual i at cycle t
          #c.W:     Cost of well state due to continued medication
          
          c.it <- 0
          c.it[M.it == "W"] <- c.W                #Cost if well
          c.it[M.it == "MI"] <- c.MI                #Cost if MI
          c.it[M.it == "ST"] <- c.ST                #Cost if stroke
          c.it[M.it == "postMI"] <- c.postMI                #Cost if MI
          c.it[M.it == "postST"] <- c.postST                #Cost if stroke
          c.it[M.it == "cD"] <- c.cD                #Cost if CVD death
          c.it[M.it == "D"] <- c.D                #Cost if dead
          return(c.it)                            #Returning costs
}

#Util function estimates the QALY at every cycle
Util <- function (M.it) {
          #M.it:    Health state occupied by individual i at cycle t
          
          u.it <- 0
          u.it[M.it == "W"] <- u.W                #QALY if well
          u.it[M.it == "MI"] <- u.MI                #QALY if MI
          u.it[M.it == "ST"] <- u.ST                #QALY if stroke
          u.it[M.it == "postMI"] <- u.postMI                #QALY if MI
          u.it[M.it == "postST"] <- u.postST                #QALY if stroke
          u.it[M.it == "cD"] <- u.cD                #QALY if CVD death
          u.it[M.it == "D"] <- u.D                #QALY if dead
          QALY <- u.it * cl
          return(QALY)                            #Returning costs
}
```
# Running the simulations
```{r,echo=FALSE}
#### Running the simulation
sim_control <- MicroSim(p.W_CVD_NT,c.W_NT)
sim_treat <- MicroSim(p.W_CVD_T,c.W_T)

#Calculating the comparision between the drugs
v.TS <-
v.CVD <- c(sim_control$count_CVD,sim_treat$count_CVD)       #CVD Events
v.C <- c(sim_control$tc_hat,sim_treat$tc_hat)               #Cost per person
v.U <- c(sim_control$tu_hat,sim_treat$tu_hat)               #QALY per person
v.cD <-c(sim_control$count_cD,sim_treat$count_cD)           #CVD Deaths
v.D <- c(sim_control$count_D,sim_treat$count_D)             #Deaths

ICER <- (v.C[2]-v.C[1])/(v.U[2]-v.U[1])                     #Incremental Cost Effectiveness Ratio

#Data Presentation
table_microsim <- data.frame(
          v.CVD,                      #CVD event
          v.D,                        #Non-CVD Deaths
          v.cD,                       #CVD deaths
          round(v.C, 0),              # costs per arm
          v.U,                        # health outcomes per arm
          c("", round(ICER, 3))       # ICER
)
rownames(table_microsim) = v.Str  # name the rows
colnames(table_microsim) = c("CVD Events","Non-CVD Deaths","CVD Deaths","Costs($)", "QALYs", "ICER") # name the columns
table_microsim                    # print the table 
```
According to WHO, if the ICER is less than thrice the GDP per Capita, the investment in the intervention should be pursued. There are alternative aproaches as well, but I haven't explored them yet.

# Diesease transition when No Treatment is provided
```{r,echo=FALSE}
sim_control[["m.CS"]]
```
# Diesease transition when Treatment is provided
```{r,echo=FALSE}
sim_treat[["m.CS"]]
```
# Plotting the curves for Survival and CVD Deaths
```{r,echo=FALSE}
xdata <- c(0:n.t+1)

cs_nt_well <- t(as.numeric(sim_control$m.CS[1,]))
cs_nt_cvddeath <- t(as.numeric(sim_control$m.CS[2,]))
cs_nt_death <- t(as.numeric(sim_control$m.CS[3,]))
cs_t_well <- t(as.numeric(sim_treat$m.CS[1,]))
cs_t_cvddeath <- t(as.numeric(sim_treat$m.CS[2,]))
cs_t_death <- t(as.numeric(sim_treat$m.CS[3,]))
```
Plot representing the subjects in well state with each cycle transition with "Blue" for Treatment cohort, and Red for No Treatment cohort 
```{r,echo=FALSE}
plot(xdata, cs_t_well, type="o", col="blue", pch="o", lty=1, ylim=c(500,1000),xlab="Cycle #", ylab="Number of Subjects" ) + points(xdata, cs_nt_well, col="red", pch="*") + lines(xdata, cs_nt_well, col="red", lty=2) 
```
Plot representing the subjects lost due to CVD deaths with each cycle transition with "Blue" for Treatment chort, and Red for No Treatment cohort 
```{r,echo=FALSE}
plot(xdata, cs_t_cvddeath, type="o", col="blue", pch="o", lty=1, ylim=c(0,120), xlab="Cycle #", ylab="Number of Subjects") + points(xdata, cs_nt_cvddeath, col="red", pch="*") + lines(xdata, cs_nt_cvddeath, col="red",lty=2)
```
```{r,echo=FALSE}
end_time <- Sys.time()
```
```{r,echo=TRUE}
time_taken <- end_time - start_time
time_taken
```
Limitations of the current model: </br>
1. The transition matrix probabilities need to be refined - We can have a detailed discussion on this </br>
2. The cohort should ideally be run till end of age (ie. death or age=80+ years), but I need to modify the p.W_CVD into a dynamic formula which calculates value based on the risk factor parameters for the individual </br>
3. The model assumes subjects to eternally stay in MI or Stroke state, and doesn't consider movement back to Well state.